Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-14 WholePlotTrajectories.Rmd
results

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011]. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments (Table \@ref(tab:Tab1)).
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009].

@@

For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
After intense disturbance however, the taxonomic richness followed a unimodal decrease with a return to pre-disturbance values.
The taxonomic evenness (Shannon and Simpson diversities) increased after disturbance, whenever its intensity.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.

@@

The impact of disturbance was examined specifically at 10, 20 and 30 years after disturbance through the linear correlation between the initial percentage of AGB removed and the value of communities Simpson and Rao diversities (both diversities of order 2).

@@

plotIDH(Data=Sim,AgbLoss=AGBloss)
plotIDH(Data=Rao,AgbLoss=AGBloss)
mtext("initial %AGB lost",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")

>>>>

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011]. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments (Table \@ref(tab:Tab1)).
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009]. 
The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found).
Biomass measurements were performed with the BIOMASS package R package [@Biomass2018].

@@

For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
After intense disturbance however, the taxonomic richness followed a unimodal decrease with a return to pre-disturbance values.
The taxonomic evenness (Shannon and Simpson diversities) increased after disturbance, whenever its intensity.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo), suggesting a similar be much delayed unimodal trajectory.

@@

The impact of disturbance was examined specifically at three times after disturbance (10, 20 and 30 years) through the linear correlation between the disturbance intensity (intial %AGB removed) and the value of communities Simpson and Rao diversities (diversities of order 2) \@ref(fig:IDHplot).
The Simpson diversity was only weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied and the impact of disturbance was observed 20 years after disturbance.

@@

plotIDH(Data=Sim,AgbLoss=AGBloss)
plotIDH(Data=Rao,AgbLoss=AGBloss)
mtext("initial AGB lost (%)",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")

### 2018-05-14 WholePlotTrajectories.tex
results

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
lowland tropical rain forest corresponding to a tropical wet climate
with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
(30-y period) with a 3-month dry season (\textless{} 100
mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
one-month dry season in March \citep{Wagner2011}. Elevation ranges
between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
to thin acrisols over a layer of transformed saprolite with low
permeability generating lateral drainage during heavy rains. The
experiment corresponds to a network of twelve 6.25ha plots that have
undergone a gradient of three logging, thinning and fuelwood treatments
(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
to a randomized plot design with three replicate blocks of four plots.
The disturbance corresponds to averages of 10 trees removed per hectare
with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
thinning of trees by poison girdling \citep{Blanc2009}.

@@

For undisturbed plots the Richness, Shannon and Simpson diversities
remained stable over the 30 years monitored. After low disturbance
intensity the richness increased, reaching a maximum gain of 14
botanical genera (plot 3 from treatment 2). After intense disturbance
however, the taxonomic richness followed a unimodal decrease with a
return to pre-disturbance values. The taxonomic evenness (Shannon and
Simpson diversities) increased after disturbance, whenever its
intensity. The evenness followed a unimodal trajectory with a just
beginning return towards initial values and a maximum value, reached
around 20 years, positively correlated to the disturbance intensity
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
were still increasing 30 years after disturbance \ref{fig:DivTaxo}),
suggesting a similar be much delayed unimodal trajectory.

@@

The impact of disturbance was examined more accurately at 10, 20 and 30
years after disturbance through the linear correlation between the
initial percentage of AGB removed and the value of communities Simpson
and Rao diversities (both diversities of order 2).

>>>>

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
lowland tropical rain forest corresponding to a tropical wet climate
with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
(30-y period) with a 3-month dry season (\textless{} 100
mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
one-month dry season in March \citep{Wagner2011}. Elevation ranges
between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
to thin acrisols over a layer of transformed saprolite with low
permeability generating lateral drainage during heavy rains. The
experiment corresponds to a network of twelve 6.25ha plots that have
undergone a gradient of three logging, thinning and fuelwood treatments
(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
to a randomized plot design with three replicate blocks of four plots.
The disturbance corresponds to averages of 10 trees removed per hectare
with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
thinning of trees by poison girdling \citep{Blanc2009}. The disturbance
intensity was measured as the percentage of aboveground biomass (\%AGB)
lost between the first inventory in 1984 and five years after
disturbance (ref to be found). Biomass measurements were performed with
the BIOMASS package R package \citep{Biomass2018}.

@@

For undisturbed plots the Richness, Shannon and Simpson diversities
remained stable over the 30 years monitored. After low disturbance
intensity the richness increased, reaching a maximum gain of 14
botanical genera (plot 3 from treatment 2). After intense disturbance
however, the taxonomic richness followed a unimodal decrease with a
return to pre-disturbance values. The taxonomic evenness (Shannon and
Simpson diversities) increased after disturbance, whenever its
intensity. The evenness followed a unimodal trajectory with a just
beginning return towards initial values and a maximum value, reached
around 20 years, positively correlated to the disturbance intensity
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
were still increasing 30 years after disturbance \ref{fig:DivTaxo},
suggesting a similar be much delayed unimodal trajectory.

@@

The impact of disturbance was examined specifically at three times after
disturbance (10, 20 and 30 years) through the linear correlation between
the disturbance intensity (intial \%AGB removed) and the value of
communities Simpson and Rao diversities (diversities of order 2)
\ref{fig:IDHplot}. The Simpson diversity was only weakly related to the
disturbance intensity (\(R^2<0.25\)) and only from 20 years after
disturbance. The Rao diversity was more strongly correlated to the
disturbance intensity (\(0.55<R^2<0.75\)) for the three times studied
and the impact of disturbance was observed 20 years after disturbance.
